Synthesising Concise Termination and Non-Termination Proofs for Bit-Vector Programs

### 2014-07-08 papers/termination/paper.tex
P --> S for nested loops
Time bound

@@

 By Theorem~\ref{thm:l-term}, if $L$ terminates then there exists a termination proof $P_T$ expressible
 in $\mathcal{L}$.  Since $\phi$ is a termination specification, $P_T \models \phi$ and so
 for any $P_N$, $(P_T, P_N) \models \phi \vee \psi$.

 Similarly if $L$ does not terminate for some input, by Theorem~\ref{thm:l-nonterm} there is a non-termination
 proof $P_N$ expressible in $\mathcal{L}$, so $P_N \models \psi$ and hence for any $P_T$, $(P_T, P_N) \models \phi \vee \psi$.

 So in either case ($L$ terminates or does not), there is a witness in $\mathcal{L}$ satisfying $\phi \vee \psi$.
\end{proof}

@@

\begin{proof}
 The background logic for Algorithm~\ref{alg:termination} is \newC, for which safety is decidable.  Also, the set of
 inputs is just the state space of $L$, which is finite.  Therefore Theorems~\ref{thm:generalised-sat} and \ref{thm:synth-semicomplete}
 hold -- the specification \lstinline!spec! is satisfiable and the instanciation of Algorithm~\ref{alg:cegis} is semi-complete.
 Therefore the call to {\sc Synthesise} terminates with a witness $(P_T, P_N)$.
 If $L$ terminates then $P_T$ is a termination proof (Theorem~\ref{thm:ct}).  Otherwise $P_N$ is a non-termination proof
 (Theorem~\ref{thm:snt}).  Exactly one of these cases holds, and the background theory being decidable means that each
 call to \lstinline|decide()| will terminate, which allows us to determine which of $P_T$ and $P_N$ is a correct proof.
\end{proof}

Hence our procedure is sound and complete for bitvector termination.

>>>>

\begin{corollary}
\label{cor:cegis-newc}
 Since safety of \newC programs is decidable, Algorithm~\ref{alg:cegis} is semi-complete
 when instanciated with \newC as a background theory.
\end{corollary}

@@

 By Theorem~\ref{thm:l-term}, if $L$ terminates then there exists a termination proof $P_T$ expressible
 in $\mathcal{L}$.  Since $\phi$ is an instance of {\bf [CT]}, $P_T \models \phi$ (Theorem~\ref{thm:ct}) and
 for any $P_N$, $(P_T, P_N) \models \phi \vee \psi$.

 Similarly if $L$ does not terminate for some input, by Theorem~\ref{thm:l-nonterm} there is a non-termination
 proof $P_N$ expressible in $\mathcal{L}$.  $\psi$ is an instance of {\bf [SNT]} and so $P_N \models \psi$
 (Theorem~\ref{thm:snt}), hence for any $P_T$, $(P_T, P_N) \models \phi \vee \psi$.

 So in either case ($L$ terminates or does not), there is a witness in $\mathcal{L}$ satisfying
 formula {\bf [GT]}: $\phi \vee \psi$.
\end{proof}

@@

\begin{proof}
 By Theorem~\ref{thm:generalised-sat}, the specification \lstinline!spec! is satisfiable.  Corollary~\ref{cor:cegis-newc}
 shows that Algorithm~\ref{alg:cegis} is semi-complete, and so the call to {\sc Synthesise} terminates with a witness $(P_T, P_N)$.
 If $L$ terminates then $P_T$ is a termination proof (Theorem~\ref{thm:ct}).  Otherwise $P_N$ is a non-termination proof
 (Theorem~\ref{thm:snt}).  Exactly one of these cases holds, and the background theory being decidable means that each
 call to \lstinline|decide()| will terminate, which allows us to determine which of $P_T$ and $P_N$ is a correct proof.
\end{proof}

Hence Algorithm~\ref{alg:termination} is sound and complete for bit-vector termination.

Theorem~\ref{thm:synth-semicomplete} provides a time bound -- the CEGIS loop can iterate at most
$2^bk$ times and each iteration requires solving 2 NP-complete problems of size < $2^bk \times \| L\|$
with $\| L \| $ the size of the program-under-analysis computing $\| L \|$.  So the
runtime is bounded by $2^{bk+1} NP(2^bk \| L \|)$.
